---
title: "Replication of 'Physiological Arousal and Political Beliefs'"
author: "Jack Luby"
date: "3/17/2019"
output:
  pdf_document: default
  html_document: default
citation_package: natbib
bibliography: bibliography.bib
---
## Abstract

This is a replication of the article "Physiological Arousal and Political Beliefs" (@renshon_lee_tingley_2014) using code and data available through the public Harvard Dataverse (@DVN/24318_2014). Below are replications of Figure 2, Table 1, and Figure 3. Several alterations are made to the replicated figures. Figure 2 is first presented in its original form then in a boxplot transformation in order to convey more distributional information. Two additional models, which include subject fixed effects, were added to table 1 to examine robustness. Finally, Figure 3's results were replicated using Bayesian distributions rather than frequentist. Formatting was improved for all three.

As an extension of the article's work, several interaction models, building upon the authors' work by integrating racial-group-specific treatment effects, are explored. These models suggest potentially interesting racial differences in treatment effects, but the power of these results are limited by a scarcity of observations.    

Great absract. Very thorough and clear to the reaedr what you will be doing.

\hfill\break

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
```

```{r libraries}
library(mediation)
library(memisc)
library(Zelig)
library(gplots)
library(knitr)
library(ri)
library(RItools)
library(gt)
library(RColorBrewer)
library(broom)
library(stringr)
library(rstanarm)
library(tidyverse)
library(stargazer)
library(rstan)
library(arm)
library(haven)
library(tidybayes)
```

```{r read-in data}
# Loading data in. I changed the csv read-in to read_csv for no good reason 
# other than I think it looks like a better function. Before starting I took 
# out a bunch of unnecessary information, and I needed to change the read-in 
# a bit to recognize the fact that the files were constained in my 
# midterm_dataverse_files folder.

anxiety <- read_csv("midterm_dataverse_files/anxiety.csv") 

# Made this subsetting a bit more tidyverse-y so that it would be a bit more 
# intuitive what is happening. I made sure that the filtering resulted in the 
# same subset as subset() by comparing the two with summary(). In this step we 
# filter out the relax video condition so that we only check the treatment 
# effect of the anxiety video against the neutral video.

noRelaxCond <- anxiety %>% 
  filter(anxcond3 != 0)

replication_data <- read_dta("midterm_dataverse_files/replicationdata.dta")
```

```{r figure 2 base}
# Interestingly, I couldn't find a function to more intuitively create a 95% CI 
# from means. The qt function below is a method found here:
# https://www.cyclismo.org/tutorial/R/confidence.html 
# which does create a quantile to find the upper and lower bounds. I found it 
# interesting that I had to declare degrees of freedom myself but maybe that's 
# due to a misunderstanding on my part.

There is ci() from the gmodels package, but what tou did is fine. That's how a lot of researchers end up calculating 
CIs

rep_data_summary <- replication_data %>% 
  
# Filtering out na's is important here so that we can calculate our summary
# statistics. This is obviously dangerous to simply remove observations but
# I felt like it made sense to throw out rows with any incomplete observations.

Always good to document what you do with missing data!
  
  filter(!is.na(anxcond3), !is.na(SCDBradVidManipAll_mean)) %>% 
  
# Grouping by the condition so that we can develop the necessary summary statistics
# as presented in the paper. Necessary for proper functioning summarize().
  
  group_by(anxcond3) %>% 
  
# Create the two summary statistics (mean and CI) which we need for the graphic.
# Saving as rep_data_summary as these are the only two figures we actually need.
  
  summarize(mean = mean(SCDBradVidManipAll_mean), 
            ci = qt(.95, df = (n() - 1))*sd(SCDBradVidManipAll_mean)/sqrt(n()))


# Pretty intuitively creating the graph here using the summary data above.

rep_data_summary %>% 
  
# Using ggplot, I create the graphic as presented in the paper. Anxcond3 must be
# recognized as a discrete variable to it is reverted to a factor.
  
  ggplot(aes(x = as.factor(anxcond3), y = mean)) + 

# Plot the mean points
  
  geom_point() + 

# Geom_errorbar allows us to add in the bars but interestingly it requires the
# input of exact values for the upper and lower bounds. These are simply the
# means plus/minus their respective error values.
  
  geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci, width=.05)) + 

# Set graph labels. Scale_x_discrete allows us to rename the numbered anxcond3
# variable which we set as a factor above.
  
  ylab("Skin Conductance Reactivity") +
  xlab("Video Condition") + 
  scale_x_discrete(labels = c('Relax','Neutral','Anxiety')) +
  labs(title = "Video Condition Significantly Affects Physiological Anxiety Response",
       subtitle = "As measured by skin conductance reactivity",
       caption = "Figure 2. Means of skin-conductance reactivity by video condition. 95% confidence intervals are presented \nin black. These distributions suggest that in a simple analysis of treatment group means, we find \nsignificantly different (i.e. non-overlapping CI) physiological responses.") +
    theme(plot.caption = element_text(hjust = 0))
```

Exactly what we were looking for for replicating figure 2. You did leave out the "n=sample size" labels though, which I think was one of the better aspects of their figure.

```{r figure 2 boxplot}
# Above I wanted to see how I would go about creating the published figure in
# R. Obviously that figure is not very illuminating as to the true distributions
# of reactivity to different video conditions. I decided to create a boxplot to
# better represent the data.

# Here we use the whole of the data so that we can display the data in more
# meaningful detail.

replication_data %>%

# Filter out incomplete video condition values because they are not meaningful
# to our figure. We do not need to filter out na SCDBradVidManipAll_mean values
# because they will automatically be excluded.

  filter(!is.na(anxcond3)) %>% 

# Ggplot is again our friend! We can simply set our as.factor anxcond3 (again
# a factor because otherwise ggplot will interpret as a continuous variable) 
# as our x variable with the reactivity values as our y variable.

  ggplot(aes(x = as.factor(anxcond3), y = SCDBradVidManipAll_mean)) + 

# A boxplot will allow us to better view the distribution of our data. There
# wasn't much else to meaningfully do with this figure without changing what it
# represents significantly (since we're only dealing with grouped means/medians).
# The boxplot actually displays the median of reactivity values.
# The notches of these boxes display the 95% CI for the median values of reactivity
# and provide a way of comparing the median values (which, of note, differs from the
# means approach in the original plot).
  
  geom_boxplot(notch = TRUE, notchwidth = .2, width = .1) + 

# Set labels as above. Once again scale_x_discrete lets us rename our discrete
# anxcond3 variables.
  
  ylab("Skin Conductance Reactivity") +
  xlab("Video Condition") + 
  scale_x_discrete(labels = c('Relax','Neutral','Anxiety')) + 
  labs(title = "Boxplot Analysis Paints a Far Murkier Picture",
       subtitle = "Observational distributions display significant overlap",
       caption = "Figure 2 Extension. Boxplots of skin-conductance reactivity by video condition. Box midpoints signify \ngroup medians (rather than means). Notches signify median 95% confidence intervals. Overall data \nvariability appears to be greater than one might expect from Figure 2 alone.") +
  theme(plot.caption = element_text(hjust = 0))
```

Cool use of boxplot as another way to visualize data. But why a murkier picture? What does what the boxplots show mean for the 
analysis moving forward. Don't leave your reader guessing here. Make sure they know exactly what you think the boxplots are showing.

\hfill\break

```{r setup table 1, results = "hide"}
# Setting up the linear models. These were a bit weird because the second
# linear model uses a different dependent variable from the first. 
# We use the noRelaxCond data because we are only looking at the anxiety 
# treatment versus the neutral condition.

Brief aside: Mediation analysis is attempting to see how much the effect of treatment (anxiety) is operating through the mediator (skin conductivity). So there is a direct effect (Anxiety -> Immigration Atts), but there is also an indirect effect (Anx -> Skin Cond. -> Immigration Atts). Mediation analysis tells if and how much of the effect of anxiety is operating through skin conductivity. So they show there is no direct effect of anxiety, which means that the skin conductivity from anxiety is actually what is resulting in changes, not just anxiety. It's generally accepted that emotion has an effect on political attitudes, but these authors want to show that it's not just changes in emotion, it's actually changes in emotion that result in a physiological response, and that it's this physical response that is causing attitude change.

# Lm1 is simply a linear model of treatment effect on physical reactivity

lm1 <- lm(SCDBradSelfReport1_mean ~ anxcond, data = noRelaxCond)

# Lm2 is a linear model of the effect of treatment, varied story conditions
# and physical reactivity on immigration opinions.

lm2 <- lm(immigration ~ anxcond + storycond + SCDBradSelfReport1_mean, data = noRelaxCond)

# In order to add an additional analysis of the robustness of these results,
# I added two additional models in which we use all available variables in the
# regressions.

lm1_robust_check <- lm(SCDBradSelfReport1_mean ~ anxcond + ideology + age + income + as.factor(race) + education, data = noRelaxCond)

lm2_robust_check <- lm(immigration ~ anxcond + storycond + SCDBradSelfReport1_mean + ideology + age + income + as.factor(race) + education, data = noRelaxCond)

# Stargazer lets us create a nice table here. We omit the statistics that
# are included automatically in stargazer but are not included in the
# published table. Then we just change around the labels to make everything
# look right.

star1 <- stargazer(lm1, lm2, lm1_robust_check, lm2_robust_check,
          omit.stat = c("adj.rsq","ll", "F", "ser"), 
          dep.var.labels = c("SC Reactivity", "Immigration Preferences", "SC Reactivity+", "Immigration Preferences+"),
          keep = c("anxcond", "storycond", "SCDBradSelfReport1_mean"),
          header = FALSE,
          title = "Main Results",
          add.lines = list(c("Subject Fixed Effects", rep("", 2), rep("$\\checkmark$", 2))),
          font.size = "small",
          covariate.labels = c("Anxiety Manipulation", "Story Condition", "SC Reactivity"),
          notes = c(""),
notes.append = TRUE)

note.latex <- "\\multicolumn{5}{l} {\\parbox[t]{15cm}{ \\textit{Notes:} Models (1) and (3) show the effect of the treatment (anxiety) on physiological reactivity while Models (2) and (4) show the effects of physiological reactivity on immigration preferences, controlling for the story condition. Models (3) and (4) include subject fixed effects for political ideology, age, income, race, and educational attainment. The significance of Model (4) suggests robustness to these fixed effects. $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01}} \\\\"
star1[grepl("Note",star1)] <- note.latex
```
Really nice table, excellent footnotes and formatting. Good replication of original findings. "Subject fixed effects" is not the right wording here. To include any type of fixed effect you would have to have each subject appearing in the data more than once, and that would be added to the regression model by something like factor(subjectid). What you want this row to be labeled is something like "Demographic Controls"

```{r print table 1, results = "asis"}
cat(star1)
```


```{r figure 3}
# Figure 3 - Causal Mediation Plot

# Here we set up the outcome and mediator models so that we can compare 
# treatment effects as explained through a mediator (ACME), against direct 
# effects, and against the total effect. Our mediator here is, as shown below, 
# SCDBradSelfReport1_mean, which represents the physiological reactivity 
# discussed in the paper. The paper contains a detailed description of the 
# innerworkings of this causal mediation analysis. Further, I changed the models 
# from a standard linear model to a bayesian linear model. This means that, using
# a t-distribution as a prior, our models iterate over the data, attempting to 
# maximize the likelihood of our posterior distribution and its coefficient
# estimates (our outputs). 

# Outcome Model

y <- bayesglm(immigration ~ anxcond + SCDBradSelfReport1_mean + storycond, 
              data= noRelaxCond)

# Mediator Model
m <- bayesglm(SCDBradSelfReport1_mean ~ anxcond + storycond, 
              data= noRelaxCond)

# Below we set the mediation analysis as set out by the authors. We use the 
# anxiety condition as the treatment and the physiological reactivity as our 
# mediator. Due to the dropobs option we re-fit our models using data rows shared 
# between the two models (not sure why this is necessary). The boot option sets a 
# nonparamtetric bootstrap.

# Mediation analysis allows us to look at the extent to which physiological
# anxiety is causing the effects explained in the article (more exclusionary
# immigration opinions). The article argues that anxiety induced by video
# treatments leads to more exclusionary immigration opinions, but we want
# to know the extent to which that is being caused by physiological anxiety
# activation as opposed to other factors.

# Mediation Analysis
m.out <- mediate(m, 
                 y, 
                 sims=500, 
                 treat="anxcond",
                 mediator="SCDBradSelfReport1_mean", 
                 dropobs=TRUE, boot=TRUE, 
                 conf.level=.90)

# This is sort of a janky way to make sure my labels get shown properly, but 
# that is what par(mar =) is doing. Helps to bump out the left side margin so 
# that all text is visible in the pdf.

par(mar=c(5,6,4,1)+.1)

# Plot the values set out above. I don't love the way this looks but I don't 
# have much of a good way to change it given that mediate objects cannot be 
# used in ggplot. 

plot(m.out, 
     labels=c("ACME\n(Physiological \nReactivity)", "Direct Effect \n(Anxiety)", "Total Effect")) %>%
  title(main = "Anxiety Conditions Affect Immigration Stance Through \nPhysiological Activation", 
        sub = c(
"Figure 3. Causal mediation plot. Treatment is anxiety video manipulation (compared to neutral condition), Mediator is skin 
conductance reactivity when answering immigration questions, Outcome is composite variable of immigration preferences. 
Horizontal lines represent 90% confidence intervals for estimates. Analysis was done using bayesian methodology, 
iteratively developing posterior data using model parameters."),
        cex.main = 1, font.main= 1, col.main= "black",
        cex.sub = .6, font.sub = 1, col.sub = "black")
```

The plot looks really good. Exactly what we were looking for. Really nice footnotes.

```{r interaction_effects_analysis, results = "hide"}
# For my extension, I decided to take a look at how the treatment effect
# varied by racial groups. At first I thought about trying to integrate
# random effects into the model but after learning more about random effects
# I realized that didn't make much sense for this data. First thing I recoded
# the race values so that they would be a bit cleaner in the output table and
# could be directly interpretable as factors. "Other" was input as "Other"
# because even with -1 in the models, the first alphabetical variable was
# dropped, I believe due to this: 
# https://stackoverflow.com/questions/41032858/lm-summary-not-display-all-factor-levels
# I decided to drop "other" because it isn't a particularly meaningful
# variable to include in the model.

noRelaxCond <- noRelaxCond %>% 
  mutate(race = case_when(race == 1 ~ "Caucasian",
                         race == 2 ~ "AfricanAmerican",
                         race == 3 ~ "Asian",
                         race == 4 ~ "Hispanic",
                         race == 7 ~ "-Other"))

# To avoid getting too complicated, I decided to only use race as an interaction
# term with video condition. With this, I felt like I would be able to make more
# straightforward analyses of the results. I tried including interactions with
# story condition as well but the interpretation got too complicated to be really
# worthwhile. I kept the models as lm's rather than stan_glm mostly just because
# I wanted to stay consistent with stargazer style.

inter_1 <- lm(immigration ~ anxcond + race + (anxcond:race - 1), data = noRelaxCond)

inter_2 <- lm(immigration ~ anxcond + storycond + race + (anxcond:race - 1), data = noRelaxCond)

inter_3 <- lm(immigration ~ anxcond + storycond + SCDBradSelfReport1_mean + race + (anxcond:race - 1), data = noRelaxCond)

# Stargazer options adapted from above. All of this should be fairly intuitive.
# I felt like the output was pretty interesting so I'm glad I went about my
# extension this way.

star2 <- stargazer(inter_1, inter_2, inter_3, 
          omit.stat = c("adj.rsq","ll", "F", "ser"),
          dep.var.labels = c("Immigration Preferences"),
          header = FALSE,
          title = "Grouped Results",
          covariate.labels = c("Anxiety Manipulation", "Story Condition", "SC Reactivity"))

# I decided to remove the "other" intercept as well because I felt like it was
# only cluttering the table when it wasn't really meaningful, especially when
# missing a coefficient.

star2 <- star2[-c(22:24)]

note.latex <- "\\multicolumn{4}{l} {\\parbox[t]{10cm}{ \\textit{Notes:} Models (1), (2), and (3) utilize racial interaction effects to examine the differences between treatment effects by racial group. $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01}} \\\\"
star2[grepl("Note",star2)] <- note.latex
```

Nice! You were the only one to explore interactions.

```{r extension_table, results = "asis"}
# This step is only necessary because I needed to manually remove the "-Other"
# intercept and then print.

cat(paste(star2))

# I didn't have any brilliant ideas on fixing formatting so I inserted a bunch of page breaks (sorry).
```
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break
\hfill\break

## Discussion 
  
These results suggest, interestingly, that the anxiety treatment might decrease exclusionary attitudes in hispanics, and increase those attitudes in all others. Anxiety might increase immigrant-exclusionary attitudes in non-immigrant groups while decreasing immigrant-exclusionary attitudes in populations with traditionally high rates of immigration. This analysis is speculative and is likely biased by the scarcity of hispanic subjects (only 3 study participants).

I'm not entirely sure why you're coming to that conclusion. You should reference the results in the regression table when making
claims so your readers can follow along.

Overall great job!

## References
